---
title: "Section 2 pre Final"
author: "Colin Linke"
date: "2024-06-03"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo = FALSE}
source("Code Transformations.R")
```

```{r models used}
full.model <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Semester*Cluster_current + Sex*Semester + Semester, random = ~1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML") 

full.model.reduced.interactions <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester, random = ~ 1 | BARCS_ID,data = data.file.long, na.action = na.exclude, method = "ML")

small.model <- lme(GPA ~ Cluster_current * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML")

pseudo.trajectory <- lme(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, method = "REML", na.action = na.exclude)

fm.time.slope <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Time*Cluster_current + Sex*Time + Time , random = ~ 1 + Time| BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML") 
```

```{r lmer models, because stupid package dissimilarities, echo = FALSE}
full.model.lmer <- lmer(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Semester*Cluster_current + Sex*Semester + Semester + (1 | BARCS_ID), data = data.file.long, na.action = na.exclude, REML = FALSE) 

full.model.lmer.reduced.interactions <- lmer(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester + (1 | BARCS_ID),data = data.file.long, na.action = na.exclude, REML = FALSE)
```


```{r testing fmri and fr}
pbkrtest::KRmodcomp(full.model.lmer, full.model.lmer.reduced.interactions) ## only allows lmer -.-
anova(full.model.reduced.interactions, full.model)
```

```{r Correlation structure}
full.model.AR1 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                     random = ~ 1 | BARCS_ID, correlation = corAR1(form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML" )

full.model.Unstructured <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corSymm(form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML"  )

full.model.CompSymm <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corCompSymm(form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML"  )

full.model.Toelpitz <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corARMA(p = 0, q = 3, form = ~ Time| BARCS_ID),
                     data = data.file.long, na.action = na.exclude, method = "ML"  )

full.model.Exponential <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corExp(form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML"  )

full.model.gaussian <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corGaus(form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML"  )

full.model.MA1 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corARMA(q = 1, form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML" )

full.model.ARMA11 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude , method = "ML"  )

full.model.ARMA12 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corARMA(p = 1, q = 2, form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML"   )

full.model.ARMA21 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corARMA(p = 2, q = 1, form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude , method = "ML" )

full.model.ARMA22 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corARMA(p = 2, q = 2, form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude , method = "ML")

AIC_values <- AIC(full.model.AR1, full.model.Unstructured, full.model.CompSymm, full.model.Toelpitz, full.model.Exponential, full.model.gaussian, full.model.MA1, full.model.ARMA11, full.model.ARMA12, full.model.ARMA21, full.model.ARMA22)
BIC_values <- BIC(full.model.AR1, full.model.Unstructured, full.model.CompSymm, full.model.Toelpitz, full.model.Exponential, full.model.gaussian, full.model.MA1, full.model.ARMA11, full.model.ARMA12, full.model.ARMA21, full.model.ARMA22)
df_scores_Cor <- data.frame(AIC_values, BIC_values)
df_scores_Cor[,-3]
```

```{r "entirely estimated DO NOT RUN", echo = FALSE}
# fmri.nlme.entirelyestimated <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
#                      random = ~ 1 | BARCS_ID, correlation = corSymm(form = ~ Time| BARCS_ID), 
#                      weights = varIdent(form = ~ 1 | BARCS_ID), 
#                      data = data.file.long, na.action = na.exclude, control = list(opt = "optim", msMaxIter = 200))
# AIC(fmri.nlme.entirelyestimated)
```


```{r plot of the residual structure}
residuals.AR1 <- plot(full.model.AR1, main = "AR1")
residuals.Unstructured <- plot(full.model.Unstructured, main = "Unstructured")
residuals.CompSymm <- plot(full.model.CompSymm, main = "Comp Symmetry")
residuals.Toelpitz <- plot(full.model.Toelpitz, main = "Toelpitz")
residuals.Exponential <- plot(full.model.Exponential, main = "Exponential") 
residuals.Gaussian <- plot(full.model.gaussian, main = "Gaussian")
residuals.MA1 <- plot(full.model.MA1, main = "MA1")
residuals.ARMA11 <- plot(full.model.ARMA11, main = "ARMA11")
residuals.ARMA12 <- plot(full.model.ARMA12, main = "ARMA12")
residuals.ARMA21 <- plot(full.model.ARMA21, main = "ARMA21")
residuals.ARMA22 <- plot(full.model.ARMA22, main = "ARMA22")

cowplot::plot_grid(residuals.AR1, residuals.Toelpitz, residuals.ARMA11, residuals.ARMA12, nrow = 2)
cowplot::plot_grid(residuals.Unstructured, residuals.CompSymm, residuals.Exponential, residuals.Gaussian)
```
flat optimization plane + similarness of the structure between ARMA 1,1 and Toelpitz --> ARMA 1,1 (less issues computationally, only 2 parameters have to be estimated instead of 3 with Toelpitz)

choosing the ARMA 1,1 structure 


```{r estimating using REML estimation method + correlation structure}
small.model.ARMA11 <- lme(GPA ~ Cluster_current * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML",  correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))

pseudo.trajectory.ARMA11 <- lme(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, method = "ML", na.action = na.exclude, correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))

time.slope.ARMA11 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Time*Cluster_current + Sex*Time + Time , random = ~ 1 + Time| BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID),control = list(maxIter = 1000, tolerance = 1e-3, opt = "optim")) 

small.model.AR1 <- lme(GPA ~ Cluster_current * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML", correlation = corAR1(form = ~ Time| BARCS_ID))

pseudo.trajectory.AR1 <- lme(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, method = "ML", na.action = na.exclude, correlation = corAR1(form = ~ Time| BARCS_ID))

time.slope.AR1 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Time*Cluster_current + Sex*Time + Time , random = ~ 1 + Time| BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML", correlation = corAR1(form = ~ Time| BARCS_ID)) 

AIC_scores <- AIC(full.model.ARMA11, full.model.AR1, small.model.ARMA11, small.model.AR1, pseudo.trajectory.ARMA11, pseudo.trajectory.AR1, time.slope.ARMA11, time.slope.AR1)
BIC_scores <- BIC(full.model.ARMA11, full.model.AR1, small.model.ARMA11, small.model.AR1, pseudo.trajectory.ARMA11, pseudo.trajectory.AR1, time.slope.ARMA11, time.slope.AR1)
df_scores <- data.frame(AIC_scores, BIC_scores)
df_scores[,-3]
```

```{r, echo = FALSE}
anova.pt <- anova(pseudo.trajectory.ARMA11)
corrected_pvalues <- (1 - (1 - c(0.001, 0.01, 0.05, 0.10))^(1/16))
colnames(anova.pt)[c(3,4)] <- c('Fvalue', 'pvalue')
anova.pt$significant <- ""
anova.pt <- anova.pt %>% mutate(significant = if_else(pvalue < corrected_pvalues[[1]], '***', 
                    if_else(pvalue < corrected_pvalues[[2]], '**', 
                            if_else(pvalue < corrected_pvalues[[3]], '*', 
                            if_else(pvalue < corrected_pvalues[[4]], '.', '')))),
                    Fvalue = round(Fvalue, digits = 4))
```
```{r 'replicating page 9 significance table'}
print(as.data.frame(anova.pt))
Anova(pseudo.trajectory.ARMA11, test.statistic = "F",  robust= "hc3", correction = "Sidak")
```

```{r, echo=FALSE}
pseudo.trajectory.lmer <- lmer(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester + (1 | BARCS_ID), data = data.file.long, REML = TRUE, na.action = na.exclude)

Anova(pseudo.trajectory.lmer, test.statistic = "F",  robust= "hc3", correction = "Sidak")
```
replicates the table with significances on page 9 




```{r estimating with REML, echo = FALSE}
full.model.ARMA11 <- lme(GPA ~ 1 + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester * Cluster_current, random = ~ 1 | BARCS_ID,data = data.file.long, na.action = na.exclude, method = "REML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))


full.model.ARMA11.slope <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester, random = ~ 1 + Time| BARCS_ID,data = data.file.long, na.action = na.exclude, method = "REML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID), control = list(maxIter = 1000, tolerance = 1e-3, opt = "optim"))
small.model.ARMA11 <- lme(GPA ~ Cluster_current * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "REML",  correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))

# time.ARMA11 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Time*Cluster_current + Sex*Time + Time , random = ~ 1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "REML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID),control = list(maxIter = 1000, tolerance = 1e-3, opt = "optim")) Not invertible. Therefore a RE time slope is included for the Time model
```
```{r}
anova(full.model.ARMA11, full.model.ARMA11.slope)
```
The performed Likelihood test with a pvalue of 0.7075 leads to a clear rejection of the inclusion of a time slope.



```{r, echo = FALSE}
pseudo.trajectory.ARMA11 <- lme(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, method = "REML", na.action = na.exclude, correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))

time.slope.ARMA11 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Time*Cluster_current + Sex*Time + Time , random = ~ 1 + Time| BARCS_ID, data = data.file.long, na.action = na.exclude, method = "REML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID),control = list(maxIter = 1000, tolerance = 1e-3, opt = "optim")) 

full.model.AR1 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester, random = ~ 1 | BARCS_ID,data = data.file.long, na.action = na.exclude, method = "REML", correlation = corAR1(form = ~ Time| BARCS_ID))

small.model.AR1 <- lme(GPA ~ Cluster_current * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "REML", correlation = corAR1(form = ~ Time| BARCS_ID))

pseudo.trajectory.AR1 <- lme(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, method = "REML", na.action = na.exclude, correlation = corAR1(form = ~ Time| BARCS_ID))

time.slope.AR1 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Time*Cluster_current + Sex*Time + Time , random = ~ 1 + Time| BARCS_ID, data = data.file.long, na.action = na.exclude, method = "REML", correlation = corAR1(form = ~ Time| BARCS_ID)) 
```


page 7

```{r}
plot(effect("Semester*Cluster_current", full.model.ARMA11, robust= "hc3", correction = "Sidak"))
plot(effect("Cluster_current*Semester", small.model.ARMA11, robust= "hc3", correction = "Sidak"))
plot(effect("Cluster_current*Time", time.slope.ARMA11, robust= "hc3", correction = "Sidak"))
```



```{r Cluster Transition plot recreation}
plot(effect("Group_transition1", pseudo.trajectory.ARMA11, robust= "hc3", correction = "Sidak"))
```

```{r Effect of Sex and Smoking on GPA}
plot(effect("Sex", full.model.ARMA11, robust= "hc3", correction = "Sidak"), main = "Effect of Gender")
plot(effect("Fager4_binary", full.model.ARMA11, robust= "hc3", correction = "Sidak"), main = "Effect of Smoking")
```

```{r assesment FM residuals}
plot(fitted(full.model.ARMA11), resid(full.model.ARMA11), 
     xlab = "Fitted Values", 
     ylab = "Residuals",
     main = "Residuals vs Fitted Plot")
abline(h = 0, col = "red")
qqPlot(resid(full.model.ARMA11), 
       main = "Normal Q-Q Plot of Residuals")

sqrt_abs_resid <- sqrt(abs(resid(full.model.ARMA11)))
plot(fitted(full.model.ARMA11), sqrt_abs_resid, 
     xlab = "Fitted Values", 
     ylab = "Square Root of |Residuals|",
     main = "Scale-Location Plot")

hist(resid(full.model.ARMA11), 
     main = "Histogram of Residuals", 
     xlab = "Residuals", 
     breaks = 10, 
     col = "gray")

plot(ranef(full.model.ARMA11), 
       main = "Random Intercept")
dim(ranef(full.model.ARMA11))
print(range(ranef(full.model.ARMA11)))
print(r.squaredGLMM(full.model.ARMA11)) ####TERRRIBLE!!!!!!!!!!!!
```

```{r assesment SM residuals}
plot(fitted(small.model.ARMA11), resid(small.model.ARMA11), 
     xlab = "Fitted Values", 
     ylab = "Residuals",
     main = "Residuals vs Fitted Plot")
abline(h = 0, col = "red")
qqPlot(resid(small.model.ARMA11), 
       main = "Normal Q-Q Plot of Residuals")

sqrt_abs_resid <- sqrt(abs(resid(small.model.ARMA11)))
plot(fitted(small.model.ARMA11), sqrt_abs_resid, 
     xlab = "Fitted Values", 
     ylab = "Square Root of |Residuals|",
     main = "Scale-Location Plot")

hist(resid(small.model.ARMA11), 
     main = "Histogram of Residuals", 
     xlab = "Residuals", 
     breaks = 10, 
     col = "gray")



plot(ranef(small.model.ARMA11), 
       main = "Random Intercept")
print(dim(ranef(small.model.ARMA11)))
print(range(ranef(small.model.ARMA11)))
print(r.squaredGLMM(small.model.ARMA11)) ####TERRRIBLE!!!!!!!!!!!!
```

```{r assesment PT residuals}
plot(fitted(pseudo.trajectory.ARMA11), resid(pseudo.trajectory.ARMA11), 
     xlab = "Fitted Values", 
     ylab = "Residuals",
     main = "Residuals vs Fitted Plot")
abline(h = 0, col = "red")
qqPlot(resid(pseudo.trajectory.ARMA11), 
       main = "Normal Q-Q Plot of Residuals")

sqrt_abs_resid <- sqrt(abs(resid(pseudo.trajectory.ARMA11)))
plot(fitted(pseudo.trajectory.ARMA11), sqrt_abs_resid, 
     xlab = "Fitted Values", 
     ylab = "Square Root of |Residuals|",
     main = "Scale-Location Plot")

hist(resid(pseudo.trajectory.ARMA11), 
     main = "Histogram of Residuals", 
     xlab = "Residuals", 
     breaks = 10, 
     col = "gray")

plot(ranef(pseudo.trajectory.ARMA11), 
       main = "Random Intercept")
dim(ranef(pseudo.trajectory.ARMA11))
print(range(ranef(pseudo.trajectory.ARMA11)))
print(r.squaredGLMM(pseudo.trajectory.ARMA11)) ### somewhat better
```
the pseudo trajectory is doing a lot better here.



```{r test RE / pooling model}
full.model.gls <- gls(GPA ~ 1 + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester * Cluster_current, data = data.file.long, na.action = na.exclude, method = "REML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))
anova(full.model.gls, full.model.ARMA11)

small.model.gls <- gls(GPA ~ Semester * Cluster_current, data = data.file.long, na.action = na.exclude, method = "REML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))
anova(small.model.ARMA11, small.model.gls)

pseudo.trajectory.gls <- gls(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester, data = data.file.long, na.action = na.exclude, correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))
anova(pseudo.trajectory.ARMA11, pseudo.trajectory.gls)
```
RIP lol. PT isn't as clear probably due to the fact that the Random effects actually do something 

```{r Multicollinearity}
vif(full.model.gls)
vif(pseudo.trajectory.gls)
```
Semester + Interaction have high / very high multicollinearity, the other fixed effects seem okay.
For the pt, the Semester and Cluster categorization in Semester are highly multicollinear, as well as the group transition variable having a high collinear value. This makes of course sense as they are related.


```{r further tests}

(ACF.pt <- ACF(pseudo.trajectory.ARMA11, maxLag=4))
plot(ACF.pt, alpha=0.01) 

#stat.ethz.ch/R-manual/R-devel/library/nlme/html/ACF.lme.html
```
expected. There is an ARMA(1,1) Cov structure in the model assumption, a lag of 1 should be showcassed here and it is.



```{r Normality / Homoscedasticity}
shapiro.test(resid(full.model.ARMA11))
shapiro.test(resid(small.model.ARMA11))
shapiro.test(resid(pseudo.trajectory.ARMA11))
shapiro.test(resid(time.slope.ARMA11)) ## additional tests are somewhat trivial
```
strong evidence that the data is not normally distributed 