---
title: "Section 2 pre Final"
author: "Colin Linke"
date: "2024-06-03"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo = FALSE, include = FALSE}
source("Code Transformations.R")
```

```{r models used}
full.model <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Semester*Cluster_current + Sex*Semester + Semester, random = ~1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML") 

full.model.reduced.interactions <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester, random = ~ 1 | BARCS_ID,data = data.file.long, na.action = na.exclude, method = "ML")

small.model <- lme(GPA ~ Cluster_current * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML")

pseudo.trajectory <- lme(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, method = "REML", na.action = na.exclude)

fm.time.slope <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Time*Cluster_current + Sex*Time + Time , random = ~ 1 + Time| BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML") 
```

```{r lmer models, because stupid package dissimilarities, echo = FALSE}
full.model.lmer <- lmer(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Semester*Cluster_current + Sex*Semester + Semester + (1 | BARCS_ID), data = data.file.long, na.action = na.exclude, REML = FALSE) 

full.model.lmer.reduced.interactions <- lmer(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester + (1 | BARCS_ID),data = data.file.long, na.action = na.exclude, REML = FALSE)
```


```{r testing fmri and fr}
pbkrtest::KRmodcomp(full.model.lmer, full.model.lmer.reduced.interactions) ## only allows lmer -.-
anova(full.model.reduced.interactions, full.model)
```

```{r Correlation structure}
full.model.AR1 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                     random = ~ 1 | BARCS_ID, correlation = corAR1(form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML" )  ##N = 3361

full.model.Unstructured <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corSymm(form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML"  ) ## N = 3361

full.model.CompSymm <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corCompSymm(form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML"  ) ## N = 3361

full.model.Toelpitz <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corARMA(p = 0, q = 3, form = ~ Time| BARCS_ID),
                     data = data.file.long, na.action = na.exclude, method = "ML"  ) ## N = 3361

full.model.Exponential <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corExp(form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML"  ) ## N = 3361

full.model.gaussian <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corGaus(form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML"  ) ## N = 3361

full.model.MA1 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corARMA(q = 1, form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML" ) ## N = 3361

full.model.ARMA11 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude , method = "ML"  ) ## N = 3361

full.model.ARMA12 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corARMA(p = 1, q = 2, form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude, method = "ML"   ) ## N = 3361

full.model.ARMA21 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corARMA(p = 2, q = 1, form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude , method = "ML" ) ## N = 3361

full.model.ARMA22 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
                       random = ~ 1 | BARCS_ID, correlation = corARMA(p = 2, q = 2, form = ~ Time| BARCS_ID), 
                     data = data.file.long, na.action = na.exclude , method = "ML") ## N = 3361

AIC_values <- AIC(full.model.AR1, full.model.Unstructured, full.model.CompSymm, full.model.Toelpitz, full.model.Exponential, full.model.gaussian, full.model.MA1, full.model.ARMA11, full.model.ARMA12, full.model.ARMA21, full.model.ARMA22)
BIC_values <- BIC(full.model.AR1, full.model.Unstructured, full.model.CompSymm, full.model.Toelpitz, full.model.Exponential, full.model.gaussian, full.model.MA1, full.model.ARMA11, full.model.ARMA12, full.model.ARMA21, full.model.ARMA22)
df_scores_Cor <- data.frame(AIC_values, BIC_values)
df_scores_Cor[,-3]
```

stargazer command
```{r "stargazer varcov"}
stargazer(df_scores_Cor[,-3], summary = FALSE, flip = F, title = "Correlation Structures Comparison")
```

```{r "entirely estimated DO NOT RUN", echo = FALSE, include = FALSE}
  # fmri.nlme.entirelyestimated <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester,
  #                      random = ~ 1 | BARCS_ID, correlation = corSymm(form = ~ Time| BARCS_ID),
  #                      weights = varIdent(form = ~ 1 | BARCS_ID),
  #                      data = data.file.long, na.action = na.exclude, control = list(opt = "optim", msMaxIter = 200))
# AIC(fmri.nlme.entirelyestimated)
```

nicht benutzt
```{r plot of the residual structure, include = FALSE, echo=FALSE}
residuals.AR1 <- plot(full.model.AR1, main = "AR1")
residuals.Unstructured <- plot(full.model.Unstructured, main = "Unstructured")
residuals.CompSymm <- plot(full.model.CompSymm, main = "Comp Symmetry")
residuals.Toelpitz <- plot(full.model.Toelpitz, main = "Toelpitz")
residuals.Exponential <- plot(full.model.Exponential, main = "Exponential") 
residuals.Gaussian <- plot(full.model.gaussian, main = "Gaussian")
residuals.MA1 <- plot(full.model.MA1, main = "MA1")
residuals.ARMA11 <- plot(full.model.ARMA11, main = "ARMA11")
residuals.ARMA12 <- plot(full.model.ARMA12, main = "ARMA12")
residuals.ARMA21 <- plot(full.model.ARMA21, main = "ARMA21")
residuals.ARMA22 <- plot(full.model.ARMA22, main = "ARMA22")

cowplot::plot_grid(residuals.AR1, residuals.Toelpitz, residuals.ARMA11, residuals.ARMA12, nrow = 2)
cowplot::plot_grid(residuals.Unstructured, residuals.CompSymm, residuals.Exponential, residuals.Gaussian)
```
There seems to be a flat optimization plane and similar correlation structures would tend to estimate similarly. AIC --> ARMA (1,1),  BIC --> AR(1). 

choosing the ARMA (1,1) structure, due to the paper deciding on the AIC. 


```{r estimating using REML estimation method + correlation structure}
small.model.ARMA11 <- lme(GPA ~ Cluster_current * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML",  correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))

pseudo.trajectory.ARMA11 <- lme(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, method = "ML", na.action = na.exclude, correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))

time.slope.ARMA11 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Time*Cluster_current + Sex*Time + Time , random = ~ 1 + Time| BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID),control = list(maxIter = 1000, tolerance = 1e-3, opt = "optim")) 

small.model.AR1 <- lme(GPA ~ Cluster_current * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML", correlation = corAR1(form = ~ Time| BARCS_ID))

pseudo.trajectory.AR1 <- lme(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, method = "ML", na.action = na.exclude, correlation = corAR1(form = ~ Time| BARCS_ID))

time.slope.AR1 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Time*Cluster_current + Sex*Time + Time , random = ~ 1 + Time| BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML", correlation = corAR1(form = ~ Time| BARCS_ID)) 
```
This is only to compare the three (plus the alternative Time as numeric variable ) models to themselves with the ARMA11 and AR1 structures. the ARMA11 structure would be chosen for all models based on the AIC, however if it were based on the BIC, then the AR1 would be prefered, with the exception of the Small Model. 


```{r, echo = FALSE}
anova.pt <- anova(pseudo.trajectory.ARMA11)
corrected_pvalues <- (1 - (1 - c(0.001, 0.01, 0.05, 0.10))^(1/16))
colnames(anova.pt)[c(3,4)] <- c('Fvalue', 'pvalue')
#anova.pt$significant <- ""
anova.pt <- anova.pt %>% mutate(corrected.pvalue = (1 - (1 - pvalue)^(1/16)),
  significant = if_else(pvalue < corrected_pvalues[[1]], '***', 
                    if_else(pvalue < corrected_pvalues[[2]], '**', 
                            if_else(pvalue < corrected_pvalues[[3]], '*', 
                            if_else(pvalue < corrected_pvalues[[4]], '.', '')))),
                    Fvalue = round(Fvalue, digits = 4))
``` 

the first one with manually computed significance with the SIdak correction, the second regular Anova output. The extra commands dont seem to do anything really unfortunately 
```{r 'replicating page 9 significance table' }
print(as.data.frame(anova.pt))
Anova(pseudo.trajectory.ARMA11)#, test.statistic = "F",  robust= "hc3", correction = "Sidak")
summary(full.model.ARMA11, correction = "Sidak")
```

```{r, echo=FALSE}
pseudo.trajectory.lmer <- lmer(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester + (1 | BARCS_ID), data = data.file.long, REML = TRUE, na.action = na.exclude)

Anova(pseudo.trajectory.lmer, test.statistic = "F",  robust= "hc3", correction = "Sidak")
```
replicates the table with significances on page 9 




```{r estimating with REML, echo = FALSE}
full.model.ARMA11 <- lme(GPA ~ 1 + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester * Cluster_current, random = ~ 1 | BARCS_ID,data = data.file.long, na.action = na.exclude, method = "ML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))


full.model.ARMA11.slope <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester, random = ~ 1 + Time| BARCS_ID,data = data.file.long, na.action = na.exclude, method = "ML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID), control = list(maxIter = 1000, tolerance = 1e-3, opt = "optim"))

#time.ARMA11 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Time*Cluster_current + Sex*Time + Time , random = ~ 1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "ML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID),control = list(maxIter = 1000, tolerance = 1e-3, opt = "optim")) #Not invertible. Therefore a RE time slope is included for the Time model
```
```{r}
AIC(full.model.ARMA11, full.model.ARMA11.slope)
BIC(full.model.ARMA11, full.model.ARMA11.slope)
```
Based on the AIC and the BIC criterium, the inclusion of the Time variable into the random variable instead of the correlation structure is not beneficial. A model in which Time is included in both the correlation structure in addition to being a random effect isn't feasible, because in such a case the coeffecient matrix isn't invertible anymore.



```{r, echo = FALSE}
full.model.ARMA11 <- lme(GPA ~ 1 + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Cluster_current * Semester , random = ~ 1 | BARCS_ID,data = data.file.long, na.action = na.exclude, method = "REML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))

small.model.ARMA11 <- lme(GPA ~ Cluster_current * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "REML",  correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))

pseudo.trajectory.ARMA11 <- lme(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, method = "REML", na.action = na.exclude, correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))

time.slope.ARMA11 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Time*Cluster_current + Sex*Time + Time , random = ~ 1 + Time| BARCS_ID, data = data.file.long, na.action = na.exclude, method = "REML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID),control = list(maxIter = 1000, tolerance = 1e-3, opt = "optim")) 

full.model.AR1 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester*Cluster_current + Semester, random = ~ 1 | BARCS_ID,data = data.file.long, na.action = na.exclude, method = "REML", correlation = corAR1(form = ~ Time| BARCS_ID))

small.model.AR1 <- lme(GPA ~ Cluster_current * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, na.action = na.exclude, method = "REML", correlation = corAR1(form = ~ Time| BARCS_ID))

pseudo.trajectory.AR1 <- lme(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester, random = ~ 1 | BARCS_ID, data = data.file.long, method = "REML", na.action = na.exclude, correlation = corAR1(form = ~ Time| BARCS_ID))

time.slope.AR1 <- lme(GPA ~ 1 + Cluster_current + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Sex*Cluster_current + Time*Cluster_current + Sex*Time + Time , random = ~ 1 + Time| BARCS_ID, data = data.file.long, na.action = na.exclude, method = "REML", correlation = corAR1(form = ~ Time| BARCS_ID)) 
```


page 7

```{r}
plot(effect("Cluster_current*Semester", full.model.ARMA11, robust= "hc3", correction = "Sidak"), main = "Full Model: Cluster & Semester")
plot(effect("Cluster_current*Semester", small.model.ARMA11, robust= "hc3", correction = "Sidak"), main = "Small Model: Cluster & Semester")
plot(effect("Cluster_current*Time", time.slope.ARMA11, robust= "hc3", correction = "Sidak"), main = "Time Model: Cluster & Semester")
```


```{r}
testfm <- emmeans(full.model.ARMA11, ~Cluster_current*Semester | Semester)
testsm <- emmeans(small.model.ARMA11, ~Cluster_current*Semester )
print(pairs(testfm))
pairs(testsm)
```


```{r Cluster Transition plot recreation}
plot(effect("Group_transition1", pseudo.trajectory.ARMA11, robust= "hc3", correction = "Sidak"), main = "Effect Cluster Transition")
```

```{r Effect of Sex and Smoking on GPA}
plot(effect("Sex", full.model.ARMA11, robust= "hc3", correction = "Sidak"), main = "Effect of Gender")
plot(effect("Fager4_binary", full.model.ARMA11, robust= "hc3", correction = "Sidak"), main = "Effect of Smoking")
```

```{r "fitted vs residuals", echo=FALSE}
fitted.fm <- fitted(full.model.ARMA11)
resid.fm <- resid(full.model.ARMA11)
df.fm <- data.frame(fitted.fm, resid.fm)

fitted.sm <- fitted(small.model.ARMA11)
resid.sm <- resid(small.model.ARMA11)
df.sm <- data.frame(fitted.sm, resid.sm)

fitted.pt <- fitted(pseudo.trajectory.ARMA11)
resid.pt <- resid(pseudo.trajectory.ARMA11)
df.pt <- data.frame(fitted.pt, resid.pt)

ggplot.fm <- ggplot(df.fm, aes(x = fitted.fm, y = resid.fm)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  labs(title = "Residuals vs Fitted values Full Model", x = "Fitted values", y = "Residuals") +
  theme_bw()


ggplot.sm <- ggplot(df.fm, aes(x = fitted.sm, y = resid.sm)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  labs(title = "Residuals vs Fitted values Small Model", x = "Fitted values", y = "Residuals") +
  theme_bw()


ggplot.pt <- ggplot(df.pt, aes(x = fitted.pt, y = resid.pt)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  labs(title = "Residuals vs Fitted values Pseudo Trajectory", x = "Fitted values", y = "Residuals") +
  theme_bw()

ggplot.fm
ggplot.sm
ggplot.pt

shapiro.test(resid(full.model.ARMA11))
shapiro.test(resid(small.model.ARMA11))
shapiro.test(resid(pseudo.trajectory.ARMA11))
```


```{r assesment FM residuals}



plot(fitted(full.model.ARMA11), resid(full.model.ARMA11), 
     xlab = "Fitted Values", 
     ylab = "Residuals",
     main = "Residuals vs Fitted Plot")
abline(h = 0, col = "red")
qqPlot(resid(full.model.ARMA11), 
       main = "Normal Q-Q Plot of Residuals")

sqrt_abs_resid <- sqrt(abs(resid(full.model.ARMA11)))
plot(fitted(full.model.ARMA11), sqrt_abs_resid, 
     xlab = "Fitted Values", 
     ylab = "Square Root of |Residuals|",
     main = "Scale-Location Plot")

hist(resid(full.model.ARMA11), 
     main = "Histogram of Residuals", 
     xlab = "Residuals", 
     breaks = 10, 
     col = "gray")

plot(ranef(full.model.ARMA11), 
       main = "Random Intercept")
dim(ranef(full.model.ARMA11))
print(range(ranef(full.model.ARMA11)))
print(r.squaredGLMM(full.model.ARMA11)) ####TERRRIBLE!!!!!!!!!!!!
```

```{r assesment SM residuals}
plot(fitted(small.model.ARMA11), resid(small.model.ARMA11), 
     xlab = "Fitted Values", 
     ylab = "Residuals",
     main = "Residuals vs Fitted Plot")
abline(h = 0, col = "red")
qqPlot(resid(small.model.ARMA11), 
       main = "Normal Q-Q Plot of Residuals")

sqrt_abs_resid <- sqrt(abs(resid(small.model.ARMA11)))
plot(fitted(small.model.ARMA11), sqrt_abs_resid, 
     xlab = "Fitted Values", 
     ylab = "Square Root of |Residuals|",
     main = "Scale-Location Plot")

hist(resid(small.model.ARMA11), 
     main = "Histogram of Residuals", 
     xlab = "Residuals", 
     breaks = 10, 
     col = "gray")



plot(ranef(small.model.ARMA11), 
       main = "Random Intercept")
print(dim(ranef(small.model.ARMA11)))
print(range(ranef(small.model.ARMA11)))
print(r.squaredGLMM(small.model.ARMA11)) ####TERRRIBLE!!!!!!!!!!!!
```

```{r assesment PT residuals}
plot(fitted(pseudo.trajectory.ARMA11), resid(pseudo.trajectory.ARMA11), 
     xlab = "Fitted Values", 
     ylab = "Residuals",
     main = "Residuals vs Fitted Plot")
abline(h=0, col = "red")
qqPlot(resid(pseudo.trajectory.ARMA11), 
       main = "Normal Q-Q Plot of Residuals")

sqrt_abs_resid <- sqrt(abs(resid(pseudo.trajectory.ARMA11)))
plot(fitted(pseudo.trajectory.ARMA11), sqrt_abs_resid, 
     xlab = "Fitted Values", 
     ylab = "Square Root of |Residuals|",
     main = "Scale-Location Plot")

hist(resid(pseudo.trajectory.ARMA11), 
     main = "Histogram of Residuals", 
     xlab = "Residuals", 
     breaks = 10, 
     col = "gray")

plot(ranef(pseudo.trajectory.ARMA11), 
       main = "Random Intercept")
dim(ranef(pseudo.trajectory.ARMA11))
print(range(ranef(pseudo.trajectory.ARMA11)))
print(r.squaredGLMM(pseudo.trajectory.ARMA11)) ### somewhat better
```
the pseudo trajectory is doing a lot better here.



```{r test RE / pooling model}
full.model.gls <- gls(GPA ~ 1 + Sex + Age1stround + SATMath + SATVerbal + SATWriting + Fager4_binary + FH_binary + STAI_SELF_Total + BDI_SELF_Total + Parental_SES  + Semester * Cluster_current, data = data.file.long, na.action = na.exclude, method = "REML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))
anova(full.model.gls, full.model.ARMA11)

small.model.gls <- gls(GPA ~ Semester * Cluster_current, data = data.file.long, na.action = na.exclude, method = "REML", correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))
anova(small.model.gls, small.model.ARMA11)

pseudo.trajectory.gls <- gls(GPA ~ 1 + Fager4_binary + FH_binary + Sex + Cluster_SEM1 + Semester + Age1stround + SATMath + SATVerbal + SATWriting + STAI_SELF_Total + BDI_SELF_Total + Parental_SES + Cluster_SEM1 * Semester + Group_transition1 * Semester, data = data.file.long, na.action = na.exclude, correlation = corARMA(p = 1, q = 1, form = ~ Time| BARCS_ID))
anova(pseudo.trajectory.ARMA11, pseudo.trajectory.gls)
```
RIP lol. PT isn't as clear probably due to the fact that the Random effects actually do something 

```{r Multicollinearity}
vif(full.model.gls)
vif(pseudo.trajectory.gls)
```
Semester + Interaction have high / very high multicollinearity, the other fixed effects seem okay.
For the pt, the Semester and Cluster categorization in Semester are highly multicollinear, as well as the group transition variable having a high collinear value. This makes of course sense as they are related.


```{r further tests}

(ACF.pt <- ACF(pseudo.trajectory.ARMA11, maxLag=4))
plot(ACF.pt, alpha=0.01) 

ACF.fm <- ACF(full.model.ARMA11, maxLag=4)
plot(ACF.fm, alpha=0.01)

ACF.sm <- ACF(small.model.ARMA11, maxLag=4)
plot(ACF.sm, alpha=0.01)



#stat.ethz.ch/R-manual/R-devel/library/nlme/html/ACF.lme.html
```
expected. There is an ARMA(1,1) Cov structure in the model assumption, a lag of 1 should be showcassed here and it is.



```{r Normality / Homoscedasticity}
shapiro.test(resid(full.model.ARMA11))
shapiro.test(resid(small.model.ARMA11))
shapiro.test(resid(pseudo.trajectory.ARMA11))
shapiro.test(resid(time.slope.ARMA11)) ## additional tests are somewhat trivial
```
strong evidence that the data is not normally distributed 